---
title: "Root Hair Nuclear Shape"
output:   
  md_document:
    variant: markdown_github
---
#Introduction
Myosin XI-I is an unique member of the myosin XI family of unconventional molecular motor proteins in the flowering plant *Arabidopis thaliana*.  XI-I is a nucleocytoplasmic linker, connecting the outer nuclear envelope to the actin cytoskeleton.  The *xi-i* insertional knockout line *xi-iT3* exhibits a defective nuclear phenotype, with rounded, invaginated nuclei easily visible in root hair cells
The circularity indices of nuclei in 4 DAG (days after germination) wild-type and *iT3* seedlings were measured using ImageJ.  (Note: This is not my real data set, I made this one up)
```{r}
nucshape<-read.table('../data/nucdata.txt', header=TRUE)

```
##Results
```{r}
library(lattice)
library(RColorBrewer)
library(latticeExtra)
knitr::opts_chunk$set(
  fig.path = "../plots/")
```

A cumulative distribution function comparing the two data sets was made in lattice using the following code:
```{r nuc_shape_ecdf}
WT <- nucshape$WT
iT3 <- nucshape$iT3
nucshapeecdf <- ecdfplot(~ WT + iT3, data=nucshape, auto.key=list(space='right'))
print(nucshapeecdf)
```


A Kolmogov-Smirnov test was performed to compare the two distributions
```{r}
kstest<-ks.test(WT, iT3, alternative = 'two.sided')
print(kstest)
```
A p-value of 
```{r, echo=FALSE}
print(kstest$p.value)
```
suggests that there is a significant difference between these two distributions.  

##Discussion

iT3 nuclei appear to generally be rounder(i.e. have a higher circularity index) than wild-type nuclei.  This suggests that myosin XI-I plays a role in maintaining proper nuclear shape in *Arabidopsis*