---
title: "Root Hair Nuclear Shape"
output: html_notebook
---
#Introduction
Myosin XI-I is an unique member of the myosin XI family of unconventional molecular motor proteins in the flowering plant Arabidopis thaliana.  XI-I is a nucleocytoplasmic linker, connecting the outer nuclear envelope to the actin cytoskeleton.  The *xi-i* insertional knockout line *xi-iT3* exhibits a defective nuclear phenotype, with rounded, invaginated nuclei easily visible in root hair cells
The circularity indices of nuclei in 4 DAG (days after germination) wild-type and *iT3* seedlings were measured using ImageJ.  (Note: This is not my real data set, I made this one up)
```{r covars, results='asis'}
library(knitr)
nucshape<-read.table('data/nucdata.txt', header=TRUE)
kable(nucshape, align='l')
```
##Results

A cumulative distribution function comparing the two data sets was made in lattice using the following code:
```{r, echo=TRUE}
library(lattice)
library(RColorBrewer)
library(latticeExtra)
WT <- nucshape$WT
iT3 <- nucshape$iT3
ecdfplot(~ WT + iT3, data=nucshape, auto.key=list(space='right'))
```
A Kolmogov-Smirnov test was performed to compare the two distributions
```{r}
kstest<-ks.test(WT, iT3, alternative = 'two.sided')
print(kstest)
```
A p-value of 
```{r, echo=FALSE}
print(kstest$p.value)
```
suggests that there is a significant difference between these two distributions.  iT3 nuclei appear to generally be rounder(i.e. have a higher circularity index) than wild-type nuclei.